# Internet Archive API

<chauthors>Lukas Isermann</chauthors> <br><br>

```{r internet-archive-1, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache=TRUE)

# setting cache to TRUE here allows that single API calls do not have to be run every time when knitting index or the single script, but only when something has been changed in index or single script
```

You will need to install the following packages for this chapter (run the code):

```{r internet-archive-2, echo=FALSE, comment=NA}
.gen_pacman_chunk("Internet_archive")
```

## Provided services/data

-   *What data/service is provided by the API?*

The Internet Archive API [@InternetArchive2016ia] and the [archiveRetriever](<https://github.com/liserman/archiveRetriever>) [@Gavras2022ia] give access to mementos stored in the [Internet Archive](<http://web.archive.org>). The Internet Archive is a non-profit organisation which builds and provides a digital library of Internet sites and other digitally stored artefacts such as books, audio recordings, videos, images and software programs. Today, the library stores about 588 billion web pages and covers a time span of over 25 years [@InternetArchive2022ia]. The Internet Archive API offers options to search and query the Internet Archive and to retrieve information on whether a certain Url is archived and currently available in the Internet Archive. Additionally, the API allows users to retrieve the Urls and the specific time stamps of the mementos available, and gives options to limit the time frame of any search and the type of resource searched for. The R-package archiveRetriever offers easy access to the base functions of the Internet Archive API. Additionally, the archiveRetriever is designed to help with web scraping information from the Internet Archive. Besides offering access to the availability and Urls of Internet Archive mementos, the archiveRetriever allows to retrieve the Urls of the mementos of any sub pages that may be linked to in the original memento. Further, the archiveRetriever offers a function to easily scrape information from these mementos.

## Prerequesites

-   *What are the prerequisites to access the API (authentication)?*

The API is free and accessible without any authentification via HTTP Urls, httr and the [archiveRetriever](<https://github.com/liserman/archiveRetriever>).

## Simple API call

-   *What does a simple API call look like?*

The API provides searches and options that can be accessed via HTTP Urls. These Urls take the pattern:

<http://web.archive.org/cdx/search/cdx?parameters/>

As parameters, the API takes the obligatory input `url`. Further, you can refine your search by setting a timeframe for your search with the parameters `from` and `to`, and specify the match type of the Url with `matchType` as being `exact`, `prefix`, `host`, or `domain` to indicate whether you are looking for results matching the exact Url, results with the given Url as prefix, all matching results from the host archive.org, or all matching results from the host archive.org including all sub hosts *.archive.org [@InternetArchive2016ia].

Further, you can collapse search results by any of the other parameters with `collapse`. To collapse results by day you could for example use the option `collapse=timestamp:8` to collapse by time unit indicated by the 8th character of the time stamp (time stamps are stored in the format yyyymmddhhmmss).

A complete API call can then look like the example below:

<http://web.archive.org/cdx/search/cdx?url=nytimes.com&from=20191201&to=20191202&collapse=timestamp:8&matchType=exact&output=json/>

The output of the API is a json-file containing information on mementos stored in the Internet Archive. Most importantly, the json-file returns the timestamp and the original Url of the memento. This information can then be used to generate memento Urls.

A detailed description of the different functions of the API can be found on  [github](<https://github.com/internetarchive/wayback/tree/master/wayback-cdx-server#readme>) [@InternetArchive2016ia].

## API access

-   *How can we access the API from R (httr + other packages)?*

Instead of typing the API request into our browser, we can use the httr package's GET function to access the API from R.

```{r internet-archive-3, message = FALSE}
# Load required packages
library(httr)
library(jsonlite)
library(tibble)

# API call
res <- GET("http://web.archive.org/cdx/search/cdx?", query = list(
  url = "nytimes.com",         
  matchType = "url",
  from = "20191201",
  to = "20191202",
  collapse = "timestamp:8",
  output = "json"
         ))

# Translate json output to tibble
result <- httr::content(res, type = "text")
result <- fromJSON(result)
result <- as_tibble(result)
names(result) <- result[1,]
result <- result[-1,]

result

```

This output can also be used to generate the Urls of the stored mementos.

```{r internet-archive-4}
# Generate vector of memento Urls
urls <- paste0("https://web.archive.org/web/", result$timestamp, "/", result$original)

urls

```

### Using the archiveRetriever {-}

Alternatively, we can use the archiveRetriever to access selected functions of the Internet Archive API. The archiveRetriever offers four different functions, `archive_overview`, `retrieve_urls`, `retrieve_links`, and `scrape_urls`. While `archive_overview` and `retrieve_urls` work as a wrapper for the Internet Archive API, `retrieve_links` and `scrape_urls` go beyond the functionality of the API and simplify larger data collections and web scraping from the Internet Archive.

`archive_overview` queries the Internet Archive API to get an overview over the dates for which mementos of a specific Url are stored in the Internet Archive. The function outputs a nicely formatted calendar to graphically display the Url's availability in the Internet Archive.

```{r internet-archive-5}
library(archiveRetriever)

archive_overview("nytimes.com", startDate = "20051101", endDate = "20051130")

```

`retrieve_urls` offers direct access to the memento Urls of any Url stored in the Internet Archive. The function takes the base Url and the time frame as inputs and returns a vector of memento Urls. For convenience, `retrieve_urls` automatically collapses mementos by day. However, this can be deactivated with the option `collapseDate = FALSE`.

```{r internet-archive-6}

urls <- retrieve_urls(homepage = "nytimes.com", startDate = "20191201", endDate = "20191202", collapseDate = FALSE)

head(urls)
```

Oftentimes researchers and practitioners do not know the specific Urls they are interested in, but want to retrieve information from all or some subpages of a base Url, e.g. when scraping news content from online newspapers. For this purpose, the archiveRetriever offers the function `retrieve_links`. `retrieve_links` takes any number of memento Urls (obtained via `retrieve_urls`) as input and returns a data.frame listing all links of mementos referenced to in the original memento. This enables users e.g. to obtain all mementos of newspaper articles from the New York Times linked to on the memento of the homepage of the newspaper.

```{r internet-archive-7}

links <- retrieve_links("http://web.archive.org/web/20191201001433/https://www.nytimes.com/")

head(links)
```

Finally, memento Urls and their time stamps are seldom the information of interest for anyone using the Internet Archive. Therefore, the archiveRetriever offers the function `scrape_urls` to easily scrape any content from Internet Archive mementos obtained by the previous functions. The function takes a memento of the Internet Archive and a named vector of XPaths (or CSS) as obligatory inputs and results in a tibble with the content scraped using the XPath/CSS selectors.

```{r internet-archive-8}
nytimes_article <- scrape_urls(Urls = "http://web.archive.org/web/20201001004918/https://www.nytimes.com/2020/09/30/opinion/biden-trump-2020-debate.html",
                               Paths = c(title = "//h1[@itemprop='headline']",
                                         author = "//span[@itemprop='name']",
                                         date = "//time//text()",
                                         article = "//section[@itemprop='articleBody']//p"))

nytimes_article

```

`scrape_urls` comes with many options that facilitate the scraping of large amounts of Internet Archive mementos and make the scraping more flexible. A detailed description of the different functions of the archiveRetriever including `scrape_urls` can be found on [github](<https://github.com/liserman/archiveRetriever/#readme>) and in the [package documentation](<https://rdrr.io/cran/archiveRetriever/man/>).

## Social science examples

-   *Are there social science research examples using the API?*

The Internet Archive is still a rarely used source for social sciences. While there are some works focused directly on the Internet Archive and its contents [e.g. @Milligan2016ia; @Littman2018ia; @Hale2017ia], or the development of the internet itself [e.g. @Hale2014ia; @Brugger2017ia], research using he Internet Archive as data source for questions unrelated to the Internet or the Internet Archive itself are scarce. A notable exception is @Gavras2022, who uses the Internet Archive to access newspaper articles from a total of 86 online newspapers in 29 countries across Europe to research European media discourse. 




